### Ex 4.48

*For a simple random sample of* $n$ *subjects, explain why it is about 95% likely that the sample proportion has error no more than* $1/\sqrt{n}$ *in estimating the population proportion. (Hint: To show this "*$1/\sqrt{n}$ *rule", find two standard errors when* $\pi = 0.50$*, and explain how this compares to two standard errors at other values of* $\pi$*.) Using this result, show that* $n = 1/M^2$ *is a safe sample size for estimating a proportion to within* $M$ *with 95% confidence.*

**Solution**

TBD

simulation:
```{r}
n=100 #sample size
pp=0.7 #population proportion
mcv = c()

for (i in 1:9999) { #Monte Carlo
  sample = rbinom(n=n, size=1, prob=pp) #size 1 is Bernoulli
  err = abs(sum(sample)/n - pp)
  mcv[i] = err <= 1/sqrt(n)
}

p = sum(mcv)/length(mcv)
p #about 95%
```

draft ex solution (1st part):
  * The standard error of a statistic is the standard deviation of its sampling distribution
  (or an estimate of that standard deviation)
  * Se la statistic è la sample mean, questa ha distribuzione normale con standard deviation
  \sigma/(sqrt(n)) per il CLT, e quindi lo standard error è \sigma/(sqrt(n))
  * If we are estimating a proportion, we are sampling bernoulli random variables and the
  proportion is the probability parameter of the bernoulli distribution that is also its mean.
  * The standard deviation of a Bernoulli distribution is sqrt((1-p)*p), so the standard error is
  sqrt((1-p)*p)/sqrt(n)
  * for a normal distribution, we have about 95% probability that our sample fall between 2
  standard deviation (standard error in our case) of its distribution, that means that about 95% of
  the times, our sample is between -2*sqrt((1-p)*p)/sqrt(n) and 2*sqrt((1-p)*p)/sqrt(n).
  * also sqrt((1-p)*p)/sqrt(n) is maximized by p=0.5, so 95% prob to be between -1/sqrt(n) and
  1/sqrt(n)
  
draft ex solution (2nd part):
  * for 95% conf we must be within 1/sqrt(n) for the previous result
  * If n=1/M^2 then 1/sqrt(n) = M


### Ex 5.16

*An experiment used a sample of college students to investigate whether cell phone use impairs drivers’ reaction times. On a machine that simulated driving situations, at irregular periods a target flashed red or green. Participants were instructed to press a brake button as soon as possible when they detected a red light. Under the cell phone condition, each student carried out a conversation on a cell phone with someone in a separate room. In the control condition, the same students listened to a radio broadcast. The "CellPhone" data file records the students’ mean response times (in milliseconds) over several trials for each condition,* $\{y_{i1}\}$ *for the cell phone condition and* $\{y_{i2}\}$ *for control.*

a. *The comparisons of means or proportions in this chapter assume independent samples for the two groups. Explain why the samples for these two conditions are dependent rather than independent.*

b. *To compare* $\mu_1$ *and* $\mu_2$*, you can use* $\{d_i = y_{i1} − y_{i2},\; i=1,...,n\}$*, here with* $n=8$*. Specify the parameter* $\mu_d$ *and* $H_0$ *for doing this, and explain why* $\mu_d = \mu_1 - \mu_2$*.*

c. *State the assumptions and test statistic, explain why it has a t-distribution with* $df = n−1$*. Report the P-value with two-sided* $H_a$*, and interpret. (The test is called a matched-pairs t-test. Matched-pairs analyses also are possible with confidence intervals, as Section 4.4.3 did in comparing weights of anorexic girls before and after a period of treatment by analyzing the mean difference in weights.)*

**Solution**

#### a.

The samples for these two conditions are dependent because the students are the same.

#### b.

TBD

Being the students the same, to compare ... we can use ...

To compare the population means $\mu_1$ and $\mu_2$ of two groups, we usually test the null hypothesis $H_0: \mu_1 = \mu_2$ or equivalently $H_0: \mu_1 - \mu_2 = 0$.

But in this case we can also test $H_0: \mu_d = 0$ with $\mu_d = {1 \over 8} \sum_{i=1}^8 d_i$ since:
$$
\mu_d = {1 \over 8} \sum_{i=1}^8 d_i = {1 \over 8} \sum_{i=1}^8 (y_{i1} − y_{i2}) =
{1 \over 8} \sum_{i=1}^8 y_{i1} - {1 \over 8} \sum_{i=1}^8 y_{i2} = \mu_1 - \mu_2
$$

#### c.

TBD

* We assume that the differences $\{d_i = y_{i1} − y_{i2},\; i=1,...,8\}$ are approximately normally distributed since the mean response times $y_{i1}$ and $y_{i2}$ are approximately normally distributed for the Central Limit Theorem.

Since we want to do a matched-pairs t-test, the test statistic is:
$$
T = {\bar Y - \mu_0 \over {s \over \sqrt n}}
$$
It has a t-distribution with $df = n−1 = 7$ (instead of a standard normal distribution) because it uses the sample standard deviation $s$ (instead of the population standard deviation $\sigma$, which is unknown).

.....................


```{r}
data = read.table("http://stat4ds.rwth-aachen.de/data/CellPhone.dat", header=TRUE)
data
```
