
---
title: "Final Project Group C"
author: "G. Alessio, A. Campagnolo, M. Polo, G. Sarnelli"
date: "2023-01-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Data Exploration 

```{r}
path <- '../..' # go up two folders
filename = '/winequality-white.csv'

data = read.csv(paste(path,filename, sep = ""), sep = ';') # load data
print(data)
```
```{r}
# Explore

library(tidyverse)
library(hrbrthemes)

plot_feature_hist <- function(data, col, name, color) {
  #bw <- 2 * IQR(data[,col]) / length(data[,col])^(1/3) # Freedman-Diaconis’s Rule
  return(
    p <- data %>%
      ggplot( aes_string(x=col)) +
        geom_histogram(fill=color, color="#e9ecef", alpha=0.9) +
        ggtitle(name) +
        theme_ipsum() +
        theme(plot.title = element_text(size=12))
    )
}

cols <- names(data)
hist_colors <- c("#003f5c", "#376a35", "#2f4b7c", "#89b8ba", "#689546", "#665191", "#a05195", "#d45087", "#f95d6a", "#ff7c43", "#ffa600", "#ed3a2d")

myplot = list()
for (i in 1:12){
  myplot[[i]] <- plot_feature_hist(data, cols[i],cols[i], hist_colors[i]) + theme(plot.margin=unit(c(1,0,0,0),"cm"))
}

grid.arrange(myplot[[1]], myplot[[2]], myplot[[3]], myplot[[4]], myplot[[5]], myplot[[6]], myplot[[7]], myplot[[8]], ncol=4, nrow=2)
grid.arrange(myplot[[9]], myplot[[10]], myplot[[11]], myplot[[12]], ncol=4, nrow=2)
```



```{r}
# Set fixed seed for reproducibility
set.seed(1);

# plot_qqplot <- function(data, col, name){
#   # Generate random numbers
#   y <- col;
#   mu <- (1 / nrow(data)) * sum(y);
#   v <- sd(y);
#   x <- rnorm(nrow(data), mu, v);
#   
#   # Sort x values
#   x <- sort(x);
#   
#   # Theoretical distribution
#   x0 <- qnorm(ppoints(nrow(data)), mu, v)
#   
#   # Plot quantile-quantile plot
#   return(plot(x = x0, y = x, xlab = "Theoretical quantiles", ylab = "Observed quantiles", title=title(name)))
# }
# 
# cols <- names(data)
# for (cols in cols){
#   print(plot_qqplot(data, data[,cols], cols))
# }
cols <- names(data)
print(cols)
for (col in cols){
  qqnorm(data[,col])
  qqline(data[,col], col = "red")
}

```
```{r}
# set seed to 0
set.seed(1)


col <- "fixed.acidity"
zscore <- (abs(data[,col]-mean(data[,col]))/sd(data[,col]))

# Check the data again. It should now have two columns: X and zscore

dim(data)
head(data)

print(zscore)

# create a new dataframe that contains only those rows 
# that have a z-score of below 3

new_data <- subset(data, zscore < 3)
dim(new_data)
# check the new dataset

data_bak <- data
```


```{r}

data <- data_bak

zscores <- data.frame()

cols <- names(data)
for (col in cols){
  zscore <- (abs(data[,col]-mean(data[,col]))/sd(data[,col]))
  zscores[,col] = zscore < 3
}
```
```{r}
cols <- names(data)
for (col in cols){
  print(col)
  print(shapiro.test(data[,col]))
}

```

```{r}
tab <- data.frame(matrix(NA, nrow = nrow(data), ncol = ncol(data)))

cols <-names(data)
#normal_cols <- c(TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE, TRUE,TRUE)
normal_cols <- c(FALSE, FALSE,FALSE, FALSE, FALSE, FALSE,FALSE, FALSE, FALSE, FALSE,FALSE, FALSE)

i=1
for (col in cols){
  if(col == "quality"){
    print("exclude quality")
    tab[,i] <- replicate(nrow(data), 0)
  } else if(normal_cols[i]==TRUE){
    tab[,i] <- (abs(data[,col]-mean(data[,col]))/sd(data[,col]))
    tab[,i] <- ifelse(tab[,i]>3,1,0)
  } else {
    data_mean <- median(data[,col])
    data_sd <- sd(data[,col])
    k <- 3
    prop_outliers <- 1/(k^2)
    lower_bound <- data_mean - k*data_sd
    upper_bound <- data_mean + k*data_sd
    tab[,i] <- ifelse(data[,col] < lower_bound | data[,col] > upper_bound,1,0)
  }
  i <- i+1
}

#summary(tab)
print(tab)
tab_rowsum <- rowSums(tab)

sum(tab_rowsum > 0)

idx_clean <- lapply(tab_rowsum, as.logical)

nrow(data)-sum(idx_clean == TRUE)
# sum rows to get a single vector

```


# Models

## Data Splitting 
- $D_{learn} + D_{test} = D$
- $D_{learn}\to D_{learn}'$ excluding the observations outside the $[-3\sigma_{x_i},+3\sigma_{x_i}]$ where $x_i$ is the observation coming from the i-feature ($i=1,\dots,11$). (or using $D_{light}$ instead of $D_{learn}$)
- Matrix of correlation $D_{learn}\to D_{light}$ for glm and gam
- ~~Assessing multicollinearity with VIFs $D_{learn}\to D_{light}$ for glm and gam~~ 

```{r}
# TRAIN TEST SPLIT
## 80% of the sample size
smp_size <- floor(0.80 * nrow(data))

## set the seed to make your partition reproducible
set.seed(1)
train_ind <- sample(seq_len(nrow(data)), size = smp_size)

train <- data[train_ind, ]
test <- data[-train_ind, ]

dim(train)
dim(test)

# CLEANING

```

## General Linear Model (GLM)

## For $D_{learn}$ and $D_{light}$

- Fit using $D_{learn}$ and $D_{learn}'$ + LASSO
- Fit using $D_{clean}$ and $D_{clean}'$ + LASSO
- ~~Leverage and cook's distance to determine a $D_{learn} \to D_{learn-out}$, $D_{clean} \to D_{clean-out}$, $D_{learn}' \to D_{learn-out}'$, $D_{clean}' \to D_{clean-out}'$~~
- ~~Fit using $D_{learn-out},D_{clean-out},D_{learn-out}',D_{clean-out}'$ + LASSO~~
- Assessing of the four ~~five~~ different fit using AIC (maybe selection for the prediction)

#### Prediction 
- Application of the model using the $D_{test}$
- Assessing of the four ~~five~~ different fit using MSE

## General Additive Model (GAM)
- Fit using $D_{learn}$ and $D_{learn}'$ + LASSO
- Fit using $D_{clean}$ and $D_{clean}'$ + LASSO
- Assessing of the four ~~five~~ different fit using AIC (maybe selection for the prediction)

#### Prediction 
- Application of the model using the $D_{test}$
- Assessing of the four ~~five~~ different fit using MSE

## Random Forest 
- Fit using $D_{learn}$
- Fit using $D_{clean}$ 

#### Prediction 
- Application of the model using the $D_{test}$
- Assessing of the two different fit using MSE

# Conclusion

#### Comparison between fit using $D_{learn}$
- 8 AIC comparison
- other $f_{comp-resps}$

#### Comparison between fit using $D_{test}$
- 10 MSE comparison
- other $f_{comp-resps}$

#### Improvements and limitations














