---
title: "GroupC_HW2"
author: "M. Polo, A. Campagnolo, S. D'Avenia, G. Cera"
date: '2022-11-10'
output:
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## FSDS - Chapter 3

### Ex 3.12

*Here the text of the first exercise.*

**Solution**

Add comments to the solution.


### Ex 3.18

*Sunshine City, which attracts primarily retired people, has 90,000 residents with a mean age of 72 years and a standard deviation of 12 years. The age distribution is skewed to the left. A random sample of 100 residents of Sunshine City has  $\bar{y} = 70$ and $s = 11$. a) Describe the center and spread of the (i) population distribution, (ii) sample data distribution. What shape does the sample data distribution probably have? Why? b)Find the center and spread of the sampling distribution of $\bar{y}$ for n = 100. What shape does it have and what does it describe? c) Explain why it would not be unusual to sample a person of age 60 in Sunshine City, but it would be highly unusual for the sample mean to be 60, for a random sample of 100 residents.*

**Solution**

#### a)

(i) The population distribution has mean 72, which we can take as a measure of center, while it has a standard deviation of 12 which we can take as a measure of spread. 

(ii) For the sample, the mean is 70 while the standard deviation is 11. Since the sample is random, we can assume that it is representative of the whole population, meaning that we can expect it will have the same distribution (more or less) as the population, i.e. a left skewed distribution. For this particular sample the mean and standard deviation are as stated above.


#### b) 

Under the reasonable assumption that a sample size of 100 is large enough for the Central Limit Theorem to hold we have that:
$$\bar{Y}_{100} \mathrel{\dot\sim} N(72, \frac{\sigma^2}{n} = 1.44)$$
Where with $\bar{Y}_{100}$ we denote the random variable describing the sample mean of a sample of size 100.\
This means that this distribution is centered at 72 and standard deviation 1.2. Under the CLT assumption this random variable has a normal distribution.\
It describes the distribution of the mean age of a sample of size 100 collected from the population of Sunshine City.


#### c)

Consider the first statement: "Sample a person of age 60 in Sunshine City" and consider $Y_i$ the random variable describing the distribution of height for a person in Sunshine City. We know that it is distributed with a mean of 72 and a standard deviation of 12, meaning that the specific value of 60 falls within 2 standard deviations of the sample mean, making it not that unlikely.

Now consider the second statement "Sample mean for a sample of size 100 is 60". We have seen in part b how the random variable $\bar{Y}_{100}$ has approximately the normal distribution with parameters stated above. Under these assumptions, the value of 60 falls exactly 10 standard deviations away from the mean, making it much less likely.



### Ex 3.28

*Here the text of the second exercise.*

**Solution**

Add comments to the solution.



## FSDS - Chapter 4 

### Ex 4.2

*Here the text of the first exercise.*

**Solution**

Add comments to the solution.


### Ex 4.4

*For the Students data file (Exercise 1.2 in Chapter 1) and corresponding population, find the ML estimate of the population proportion believing in life after death. Construct a Wald 95% confidence interval, using its formula (4.8). Interpret.*

**Solution**

Add comments to the solution.
```{r Students}
data <- read.table("http://stat4ds.rwth-aachen.de/data/Students.dat", header = TRUE)

```


### Ex 4.38

*Here the text of the second exercise.*

**Solution**

Add comments to the solution.


### Ex 4.44

*Here the text of the second exercise.*

**Solution**

Add comments to the solution.


### Ex 4.48

*Here the text of the second exercise.*

**Solution**

Add comments to the solution.



## FSDS - Chapter 5 

### Ex 5.2

*When a government does not have enough money to pay for the services that it provides, it can raise taxes or it can reduce services. When the Florida Poll asked a random sample of 1200 Floridians which they preferred, 52% (624 of the 1200) chose raise taxes and 48% chose reduce services. Let $\pi$ denote the population proportion of Floridians who would choose raising taxes. Analyze whether this is a minority of the population ($\pi <$ 0.50) or a majority ($\pi >$ 0.50) by testing H0 ∶ $\pi$ = 0.50 against Ha: $\pi \neq $ 0.50. Interpret the P-value. Is it appropriate to “accept H0? Why or why not?*

**Solution**

Add comments to the solution.

### Ex 5.12

*Here the text of the second exercise.*

**Solution**

Add comments to the solution.

### Ex 5.16

*Here the text of the second exercise.*

**Solution**

Add comments to the solution.

### Ex 5.68

*Here the text of the second exercise.*

**Solution**

Add comments to the solution.


