---
title: "GroupD_HW1"
author: "G. Marsich, M. Vicari, M. Polo, E. Malcapi"
date: "2022-10-20"
output:
  html_document:
    toc: yes
    toc_depth: '3'
    df_print: paged
  pdf_document:
    toc: yes
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## FSDS - Chapter 1

### Ex X.x

*Here the text of the first exercise.*

**Solution**

Add comments to the solution.

```{r cars, echo=T}
summary(cars)
```

### Ex X.x

*Here the text of the second exercise.*

**Solution**

Add comments to the solution.

$$ Y \sim N(0,1)$$

### Ex 1.44

### Ex 1.48

## FSDS - Chapter 2

### Ex 2.16

### Ex 2.27

### Ex 2.71

## CS - Chapter 1

### Ex 1.2

*Evaluate* $Pr(X < 0.5, Y < 0.5)$ *if* $X$ *and* $Y$ *have the following joint p.d.f.*
$$
f(x,y) = \begin{cases}
  x+3y^2/2 & 0<x<1 \;\&\; 0<y<1 \\
  0 & \text{otherwise}.
\end{cases}
$$

**The probability corresponds to the volume under the portion of interest in the graph:**

$$
\int_{0}^{1\over2} \int_{0}^{1\over2} x+{3\over2}y^2 \,dx\,dy =
\int_{0}^{1\over2} \left[ {x^2\over2} + {3\over2}y^2x \right]_0^{1\over2} \,dy =
\int_{0}^{1\over2} {1\over8}+{3\over4}y^2 \,dy =
\left[ {1\over8}y + {1\over4}y^3 \right]_0^{1\over2} = {3\over32}
$$

## CS - Chapter 3

### Ex 3.3

*Rewrite the following, replacing the loop with efficient code:*
```
n <- 100000; z <- rnorm(n)
zneg <- 0; j <- 1

for (i in 1:n) {
  if (z[i]<0) {
    zneg[j] <- z[i]
    j <- j + 1
  }
}
```
*Confirm that your rewrite is faster but gives the same result.*

**Let's time the original code:**

```{r, echo=T}
n <- 100000; z <- rnorm(n)
zneg <- 0; j <- 1

system.time( #time the original code
  for (i in 1:n) {
    if (z[i]<0) {
      zneg[j] <- z[i]
      j <- j + 1
    }
  }
)
```

**Let's time the rewritten code:**

```{r, echo=T}
system.time(
  opt_zneg <- z[z<0] #rewritten code
)
```

**Let's ensure both codes give the same result:**

```{r, echo=T}
all(zneg == opt_zneg)
```

**The rewrite is faster and gives the same result.**
