---
title: "Final Project Group C"
author: "G. Alessio, A. Campagnolo, M. Polo, G. Sarnelli"
date: "2023-01-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project aim

The goal is to model wine quality based on physicochemical tests. 

# Data Exploration

In order to obtain better results with our models we explore the original dataset coming from the following source:

```{r}
#load dataset
d_tot = read.csv("winequality-white.csv", sep=";")
```

In order to look at the data we decide to plot 12 histograms for each input features $x^{(i)}_j$ where $j=1,\dots,11$, plus $y^{(i)}=\text{"the wine-quality"}$ that is our response variable.

```{r echo=FALSE}
library(ggplot2)
library(gridExtra)

plot_feature_hist <- function(data, name, color) {
  #bw <- 2 * IQR(d_tot[,col]) / length(d_tot[,col])^(1/3) # Freedman-Diaconis’s Rule
  return(
    p <-
      ggplot(as.data.frame(data) , aes(x=data)) +
        geom_histogram(fill=color, color="#e9ecef", alpha=0.9) +
        ggtitle(name)
    )
}

cols <- names(d_tot)
hist_colors <- c("#003f5c", "#376a35", "#2f4b7c", "#89b8ba", "#689546", "#665191", "#a05195", "#d45087", "#f95d6a", "#ff7c43", "#ffa600", "#ed3a2d")

myplot = list()
for (i in 1:12){
  myplot[[i]] <- plot_feature_hist(d_tot[,cols[i]],cols[i], hist_colors[i]) + theme(plot.margin=unit(c(1,0,0,0),"cm"))
}

suppressMessages(grid.arrange(myplot[[1]], myplot[[2]], myplot[[3]], myplot[[4]], myplot[[5]], myplot[[6]], myplot[[7]], myplot[[8]], ncol=4, nrow=2))
suppressMessages(grid.arrange(myplot[[9]], myplot[[10]], myplot[[11]], myplot[[12]], ncol=2, nrow=2))
```

Coming from this preliminary look a the data most of features presents **long tail** of occurrences far from the "head" or central part of the distribution. 

Afterwards in order to address intercorrelations among two or more features in a multiple regression model (multicollinearity). We plot the Correlation Matrix:
```{r}
library(corrplot)
#plot the correlation matrix
corrplot(cor(d_tot), method = "shade")
```

While a moderate amount of multicollinearity is acceptable in a regression model, a higher multicollinearity can be a cause for concern. We can see that the **density** is highly correlated and it can be removed, as the information provided by these variables is **redundant**. 
To prove that and to consider the possibility of removing the variable we look at the VIF coefficients:
```{r message=FALSE}
library(car)
model = lm(quality ~ ., data = d_tot)
vif(model)
```

# Models

## Datasets

In order to assess the models we split the total dataset $D_{tot}$ in $D_{train} + D_{test} = D_{tot}$. We introduce the following datasets $D_{train}$, $D_{train}^{light}$, $D_{train}^{clean}$ and $D^{light,clean}_{train}$ used to obtain the models.

```{r}
#d_train (3918 rows)
set.seed(101)
train_indexes = sample(1:nrow(d_tot), size=nrow(d_tot)*0.8) #80%
d_train = d_tot[train_indexes,]
row.names(d_train) = NULL
x_train = subset(d_train, select=-quality)
y_train = d_train$quality

#d_test (980 rows)
d_test = d_tot[-train_indexes,]
row.names(d_test) = NULL
x_test = subset(d_test, select=-quality)
y_test = d_test$quality
```

1. $D_{train}$ is the dataset without any modification.

2. $D_{light}$ is the dataset obtained without the **density feature**, we remove it because the Correlation Matrix arises multicollinearity problem (more insights in the data exploration section) and in order to maintain consistency between the different models we decide to just look at it and remove the most correlated feature, as told the **density**. (*light phase*)

```{r}
#d_train_light (3918 rows)
d_train_light = subset(d_train, select=-density)
x_train_light = subset(x_train, select=-density)
y_train_light = y_train

#d_test_light (980 rows)
x_test_light = subset(x_test, select=-density)
y_test_light = y_test
```

3. $D_{clean}$ is the dataset obtained by excluding the observations outside the $[\overline{x}-3\sigma_{x_j},\overline{x}+3\sigma_{x_j}]$ where $x_j$ is the observation coming from the i-feature ($j=1,\dots,11$). (*clean phase*)

```{r}
#function to clean datasets
clean = function(dataset, x, k=4) { #4 standard deviations
  means = colMeans(x)
  sds = apply(x, MARGIN=2, FUN=sd)
  f = function(x) all(x>means-k*sds & x<means+k*sds)
  return (dataset[apply(x, MARGIN=1, FUN=f),])
}

#d_train_clean (3791 rows)
d_train_clean = clean(d_train, x=x_train)
x_train_clean = subset(d_train_clean, select=-quality)
y_train_clean = d_train_clean$quality
```

4.  $D_{light,clean}$ is the dataset obtained from $D_{light}$ by applying the *clean phase*. We decide to do the *light phase* before the *clean phase* because we don't want to remove observations coming from removing the **density feature**.

```{r}
#d_train_light_clean (still 3791 rows but could have been more)
d_train_light_clean = clean(d_train_light, x=x_train_light)
x_train_light_clean = subset(d_train_light_clean, select=-quality)
y_train_light_clean = d_train_light_clean$quality
```

Using this 4 datasets we decide to implement the following models: **GLM**, **GAM** and **RF**.

## General Linear Model (GLM)

1. Fit using $D_{train}$ + LASSO

Using the train data


```{r echo=FALSE}
library(glmnet)

# glmlasso(data,aic){} for every alla the dataset
#data=data_train
fit1 = glm(quality ~ ., family = gaussian(link = "identity"), data = d_tot)

#rescaling the data
data_scale = as.data.frame(scale(d_tot[,1:11]))
quality = d_tot$quality
data_scale = cbind(data_scale, quality)
rm(quality)

attach(data_scale)
x = cbind(fixed.acidity, volatile.acidity, citric.acid, residual.sugar, chlorides, free.sulfur.dioxide, total.sulfur.dioxide, density, pH, sulphates, alcohol)
set.seed(4)  #we fix a seed for the random operations of cv
cv = cv.glmnet(x, quality, alpha=1, family=gaussian(link = "identity"), type.measure="mse", nfolds = 10) #we do a 10 fold CV with mean squared error loss function

lambda_min = cv$lambda.min#we find the best value of lambda
fit_lasso = glmnet(x, quality, alpha=1, family=gaussian(link = "identity"), lambda = lambda_min)
#fit_lasso = glmnet(x, quality, alpha=1, family=gaussian(link = "identity"))
print(paste("The AIC of the model is ",AIC(fit1)))
#print(paste("The AIC of the model with lasso is ",AIC(fit_lasso)))

```

2. Fit using $D_{light}$ + LASSO

Using the light data

```{r echo=FALSE}
fit1 = glm(quality ~ ., family = gaussian(link = "identity"), data = d_tot)
print(paste("the AIC is ",AIC(fit1)))
```


3. Fit using $D_{clean}$ + LASSO

Using the clean data

```{r echo=FALSE}
fit1 = glm(quality ~ ., family = gaussian(link = "identity"), data = d_tot)
print(paste("the AIC is ",AIC(fit1)))
```


4. Fit using $D_{light,clean}$ + LASSO

```{r echo=FALSE}
fit1 = glm(quality ~ ., family = gaussian(link = "identity"), data = d_tot)
print(paste("the AIC is ",AIC(fit1)))
```

- Assessing of the four different fit using AIC

#### Prediction

-   Application of the model using the $D_{test}$
-   Assessing of the four different fit using MSE

## General Additive Model (GAM)

## Random Forest

- Fit using $D_{train}$

```{r}
library(randomForest) #v 4.6-14 for R 4.0.4
set.seed(101)
rf_ = randomForest(x_train, y_train, x_test, y_test, importance=TRUE)
varImpPlot(rf_, sort=TRUE)
```

```{r}
#con mtry
set.seed(101)
rf2_ = randomForest(x_train, y_train, x_test, y_test, mtry=ncol(x_train), importance=TRUE)
varImpPlot(rf2_, sort=TRUE)
```
```{r}
#con mtry e senza density
set.seed(101)
rf2_ = randomForest(x_train_light, y_train_light, x_test_light, y_test_light, mtry=ncol(x_train_light), importance=TRUE)
varImpPlot(rf2_, sort=TRUE)
```

- Fit using $D_{clean}$

#### Prediction

- Application of the model using the $D_{test}$
- Assessing of the two different fit using MSE

# Conclusion

#### Comparison between fit using $D_{train}$

- 8 AIC comparison

#### Comparison between fit using $D_{test}$

- 10 MSE comparison

#### Improvements and limitations