---
title: "Final Project Group C"
author: "G. Alessio, A. Campagnolo, M. Polo, G. Sarnelli"
date: "2023-01-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Project aim

The goal is to model wine quality based on physicochemical tests. 

# Data Exploration

In order to obtain better results with our models we explore the original dataset coming from the following source:

```{r}
#load dataset
d_tot = read.csv("winequality-white.csv", sep=";")
```

In order to look at the data we decide to plot 12 histograms for each input features $x^{(i)}_j$ where $j=1,\dots,11$, plus $y^{(i)}=\text{"the wine-quality"}$ that is our response variable.

```{r echo=FALSE}
library(ggplot2)
library(gridExtra)

plot_hist <- function(x, name, color) {
  return(ggplot(data = data.frame(x), aes(x=x)) + geom_histogram(fill=color, color="#e9ecef", alpha=0.9) + ggtitle(name))
}

hist_colors <- c("#003f5c", "#376a35", "#2f4b7c", "#89b8ba", "#689546", "#665191", "#a05195", "#d45087", "#f95d6a", "#ff7c43", "#ffa600", "#ed3a2d")

hist_list <- Map(plot_hist, data.frame(d_tot), names(d_tot), hist_colors)

suppressMessages(grid.arrange(hist_list[[1]], hist_list[[2]], hist_list[[3]], hist_list[[4]], hist_list[[5]], hist_list[[6]], ncol=3, nrow=2))
suppressMessages(grid.arrange(hist_list[[7]], hist_list[[8]], hist_list[[9]], hist_list[[10]], hist_list[[11]], hist_list[[12]],  ncol=3, nrow=2))
```

Coming from this preliminary look a the data most of features presents **long tail** of occurrences far from the "head" or central part of the distribution.

Using boxplots we can  identify presence of outliers or unusual values.

```{r}
plotboxplot <- function(x, name) {
  return(ggplot(data = data.frame(x), aes(x=x)) + geom_boxplot(width=1.4, outlier.size=1, outlier.alpha = 0.3) + ggtitle(name) + ylim(-1,1))
}

box_list <- Map(plotboxplot, data.frame(d_tot), names(d_tot))

grid.arrange(box_list[[1]], box_list[[2]], box_list[[3]], box_list[[4]], box_list[[5]], box_list[[6]], ncol=2, nrow=3)
grid.arrange(box_list[[7]], box_list[[8]], box_list[[9]], box_list[[10]], box_list[[11]], box_list[[12]], ncol=2, nrow=3)
```

Afterwards in order to address intercorrelations among two or more features in a multiple regression model (multicollinearity). We plot the Correlation Matrix:
```{r message=FALSE}
library(corrplot)
#plot the correlation matrix
corrplot(cor(d_tot), method = "shade")
```

While a moderate amount of multicollinearity is acceptable in a regression model, a higher multicollinearity can be a cause for concern. We can see that the **density** is highly correlated and it can be removed, as the information provided by these variables is **redundant**. 

# Models

## Datasets

In order to assess the models we split the total dataset $D_{tot}$ in $D_{train} + D_{test} = D_{tot}$. We introduce the following datasets $D_{train}$, $D_{train}^{light}$, $D_{train}^{clean}$ and $D^{light,clean}_{train}$ used to obtain the models.

```{r}
#d_train (3918 rows)
set.seed(101)
train_indexes = sample(1:nrow(d_tot), size=nrow(d_tot)*0.8) #80%
d_train = d_tot[train_indexes,]
row.names(d_train) = NULL
x_train = subset(d_train, select=-quality)
y_train = d_train$quality

#d_test (980 rows)
d_test = d_tot[-train_indexes,]
row.names(d_test) = NULL
x_test = subset(d_test, select=-quality)
y_test = d_test$quality
```

1. $D_{train}$ is the dataset without any modification.

2. $D_{train}^{light}$ is the dataset obtained without the **density feature**, we remove it because the Correlation Matrix arises multicollinearity problem (more insights in the data exploration section) and in order to maintain consistency between the different models we decide to just look at it and remove the most correlated feature, as told the **density**. (*light phase*)

```{r}
#d_train_light (3918 rows)
d_train_light = subset(d_train, select=-density)
x_train_light = subset(x_train, select=-density)
y_train_light = y_train

#d_test_light (980 rows)
x_test_light = subset(x_test, select=-density)
y_test_light = y_test
```

3. $D_{train}^{clean}$ is the dataset obtained by excluding the observations outside the $[\overline{x}-4\sigma_{x_j},\overline{x}+4\sigma_{x_j}]$ where $x_j$ is the observation coming from the j-feature ($j=1,\dots,11$). (*clean phase*)

```{r}
#function to clean datasets
clean = function(dataset, x, k=4) { #4 standard deviations
  means = colMeans(x)
  sds = apply(x, MARGIN=2, FUN=sd)
  f = function(x) all(x>means-k*sds & x<means+k*sds)
  return (dataset[apply(x, MARGIN=1, FUN=f),])
}

#d_train_clean (3791 rows)
d_train_clean = clean(d_train, x=x_train)
x_train_clean = subset(d_train_clean, select=-quality)
y_train_clean = d_train_clean$quality
```

4.  $D_{train}^{light,clean}$ is the dataset obtained from $D_{train}^{light}$ by applying the *clean phase*. We decide to do the *light phase* before the *clean phase* because we don't want to remove observations coming from removing the **density feature**.

```{r}
#d_train_light_clean (still 3791 rows but could have been more)
d_train_light_clean = clean(d_train_light, x=x_train_light)
x_train_light_clean = subset(d_train_light_clean, select=-quality)
y_train_light_clean = d_train_light_clean$quality
```

Using this 4 datasets we decide to implement the following models: **GLM**, **GAM** and **RF**.

## General Linear Model (GLM)
Now we analyze the 4 different datasets in order to find a linear relationship between the features of the selected dataset and the response variable **quality**.

We combine the GLM (Generalized Linear Model) with the **LASSO method** (Least Absolute Shrinkage and Selection Operator). The LASSO method introduces a penalty term that is proportional to the sum of the absolute values of the coefficients, thus it forces the model to decrease or to remove the effect of some less important feature.

In order to avoid a too simple model considering only the main effects (11 features) we consider in addition their interactions $\binom{11}{2} = 55$.
Afterwards to properly use the LASSO method we need to rescale the features of our data, so that we can avoid an imbalance between coefficients. 

Re-scaling of the $D_{train}$
```{r}
#function to scale train sets
scale_train = function(x_train) {
  x_train_int = model.matrix(as.formula( ~ .*.), x_train)[,-1]
  values = scale(x_train_int)
  return (list(
    values=values,
    means=attr(values, which="scaled:center"),
    sds=attr(values, which="scaled:scale")))
}

# scaling
x_train_s = scale_train(x_train)
```

In order to choose the smoothing parameter $\lambda$ we use **10-fold cross validation**. For each $\lambda$ value in a grid, software fits the model to a part of the data and then checks the goodness of the predictions for the **quality** in the remaining data, using the **Mean Squared Error (MSE)**. We do this 10 times, then we using the approach called **one-standard-error rule** that is select a larger $\lambda$ that has a MSE (it's a random variable) one standard error above the minimum, which moves the estimates in the direction of a greater regularization (less likely to overfit the model and more interpretability). 

Fit using $D_{train}$ scaled
```{r message=FALSE, warning=FALSE}
library(glmnet)

#function to fit glm with lasso
fit_lasso <- function(x, y){
  cv <- cv.glmnet(x, y, family=gaussian, type.measure = "mse")
  return (glmnet(x, y, family=gaussian, lambda = cv$lambda.1se))
}

# fitting
set.seed(101)
lasso = fit_lasso(x_train_s$values, y_train)
paste("The number of coefficients different from zero are",length(coef(lasso)[coef(lasso)!=0]))
coef(lasso)
```


We apply the glm trained to the test set and compute the final **MSE**. 
```{r}
# prediction
x_test_int <- model.matrix(as.formula( ~ .*.), x_test)[,-1]
x_test_s <- scale(x_test_int, center=x_train_s$means, scale=x_train_s$sds)
paste('the MSE with d_train is',assess.glmnet(lasso, newx=x_test_s, newy=y_test)$mse[1])
```

We apply successively the same procedures for the remaining 3 dataset: $D_{train}^{light}$, $D_{train}^{clean}$ and $D^{light,clean}_{train}$. Therefore:

For $D_{train}^{light}$ we have the following results:
```{r message=FALSE, warning=FALSE}
# scaling
x_train_light_s <- scale_train(x_train_light)

# fitting
set.seed(101)
lasso_light = fit_lasso(x_train_light_s$values, y_train_light)
paste("The number of coefficients different from zero are",length(coef(lasso_light)[coef(lasso_light)!=0]))

# Prediction
x_test_light_int <- model.matrix(as.formula( ~ .*.), x_test_light)[,-1]
x_test_light_s <- scale(x_test_light_int, center=x_train_light_s$means, scale=x_train_light_s$sds)
paste('the MSE with d_train_light is',assess.glmnet(lasso_light, newx=x_test_light_s, newy=y_test_light)$mse[1])

```

For $D_{train}^{clean}$ we have the following results:
```{r message=FALSE, warning=FALSE}
# scaling
x_train_clean_s = scale_train(x_train_clean)

# fitting
set.seed(101)
lasso_clean = fit_lasso(x_train_clean_s$values, y_train_clean)
paste("The number of coefficients different from zero are",length(coef(lasso_clean)[coef(lasso_clean)!=0]))

# Prediction
x_test_s2 <- scale(x_test_int, center=x_train_clean_s$means, scale=x_train_clean_s$sds)
paste('the MSE with d_train_clean is',assess.glmnet(lasso_clean, newx=x_test_s2, newy=y_test)$mse[1])
```

For $D_{train}^{light,clean}$ we have the following results:
```{r message=FALSE, warning=FALSE}
# scaling
x_train_light_clean_s = scale_train(x_train_light_clean)

# fitting
set.seed(101)
lasso_light_clean = fit_lasso(x_train_light_clean_s$values, y_train_light_clean)
paste("The number of coefficients different from zero are",length(coef(lasso_light_clean)[coef(lasso_light_clean)!=0]))

# prediction
x_test_light_s2 <- scale(x_test_light_int, center=x_train_light_clean_s$means, scale=x_train_light_clean_s$sds)
paste('the MSE with d_train_light_clean is',assess.glmnet(lasso_light_clean, newx=x_test_light_s2, newy=y_test_light)$mse[1])

```

Based on the results we have, we can conclude that the GLM fitted on the cleaned dataset performs better than the others. This seems to suggest that removing **the density feature** ($D_{train}^{light}$) cause the model to quietly under-perform and remove observations that seems to be "outliers" can slightly improve the results ($D_{train}^{clean}$).
In the other hand these results might not be significantly different. 

## General Additive Model (GAM)

#### Prediction

## Random Forest

* Two way of analyzing data:
  * **Statistical models** (eg. GLM, GAM)
    * Focus on interpretation
    * Describe relationship between variables in functional form
  * **Algorithms** (eg. decision tree)
    * Focus on prediction

\

* Decision tree (**regression tree** in our case):
  * Built by recursively splitting the data
    * Each node of the tree corresponds to a split
    * Each split is based on values of a feature
      * The one that allows the best split is selected
    * Splitting continue until a stopping criteria is met
      * Eg. min number of data in a node or max tree depth
  * Prediction made by traversing the tree until a leaf node
    * Predicted value is the mean of y of observations falling in that node

\

* **Random forest**:
  * Ensemble of decision trees (500 in our case)
  * The predictions are averaged
  * Each tree is different because:
    * built in different datasets obtained with bootstrap
    * built with some of the feature randomly dropped
  * Less interpretable than single tree but better results

\

- Fit random forest using $D_{train}$

```{r message=FALSE}
library(randomForest)
set.seed(101)

rf = randomForest(x_train, y_train, importance=TRUE)
mean((predict(rf, x_test) - y_test)^2) #mse
```

- Importance of the features

`%IncMSE`: this shows how much our model MSE increase (in percentage) if we leave out that feature

`IncNodePurity`: this is a measure of importance of a feature in determining good splits in the trees

```{r}
varImpPlot(rf)
```

- Fit using $D_{train}^{light}$

```{r}
set.seed(101)

rf_light = randomForest(x_train_light, y_train_light, importance=TRUE)
mean((predict(rf_light, x_test_light) - y_test_light)^2) #mse
```

MSE doesn't change too much

- Importance of the features

```{r}
varImpPlot(rf_light)
```

We note that the importance of `alcohol` greatly increase removing `density` (they were correlated)

- Fit using $D_{train}^{clean}$

```{r}
set.seed(101)

rf_clean = randomForest(x_train_clean, y_train_clean, importance=TRUE)
mean((predict(rf_clean, x_test) - y_test)^2) #mse
```

We note that the MSE gets slightly worse removing extreme values, meaning that probably they weren't outliers

- Importance of the features

```{r}
varImpPlot(rf_clean)
```

Same feature importance as with $D_{train}$ (we don't expect it to change since the features are the same)

- Fit using $D_{train}^{light,clean}$

```{r}
set.seed(101)

rf_light_clean = randomForest(x_train_light_clean, y_train_light_clean, importance=TRUE)
mean((predict(rf_light_clean, x_test_light) - y_test_light)^2) #mse
```

Still slightly worse results than with all the data

- Importance of the features

```{r}
varImpPlot(rf_light_clean)
```

Same feature importance as with $D_{train}^{light}$ (we don't expect it to change since the features are the same)

# Conclusions

```{r comment=Na}
library(knitr)

data = data.frame(
  GML = c(0.549, 0.544, 0.534, 0.543),
  GAM = c(0.498, 0.509, 0.830, 0.872),
  RF = c(0.383, 0.389, 0.393, 0.399),
  row.names = c("d_train", "d_train_ligh", "d_train_clean", "d_train_light_clean")
  
  
)

kable(data)
```

In our project we tried three models: GLM, GAM and Random Forest.
These three models have different limitations and strong points:
* GLM
  * highly interpretable
  * used to show the interactions between our features
* GAM
  * shows an improvement on the previous model
  * can capture the non linear effects of the features
  * It shows that all the observations carry informations relevant to the effectiveness of the models
* Random Forest
  * shows the best results among our models
  * It proved to be flexible and to be able to capture complex interactions among the features

There seems to be possible ways to improve our results.
Having more data (more samples of wine) could obviously be beneficial. We also do not exclude the possibility that there could be a greater number of significant features that could better explain the quality of the wine.
At the end we must admit that human preferences could always create a stochastic component in our model because we cannot predict subjective tastes. 




